Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-10-05 forecast/paper/supplement-text.tex
Pass over supplement
- Just formatting and copyedits
- Comment out all eval results from 2021 (but I can't seem to
  compile the Rmd because I'm missing the data files)

\section{Examining the relative advantage of using finalized rather than vintage data}

@@

\section{Aggregating with geometric mean}

In this section, we consider using the geometric mean instead of the
arithmetic mean when aggregating the weighted interval score (WIS) across
location-time pairs. 
There are three reasons why using the geometric mean may be desirable.
\begin{enumerate}
\item  WIS is right-skewed, being bounded below by zero and having
  occasional very large values.   Figure \ref{fig:wis-densities} illustrates that the
  densities appear roughly log-Gaussian.  The geometric mean is a
  natural choice in such a context since the relative ordering of forecasters is
  determined by the arithmetic mean of the {\em logarithm} of their WIS
  values.
\item In the main paper, we report the ratio of the mean WIS of a
  forecaster to the mean WIS of the baseline forecaster. Another
  choice could be to take the mean of the ratio of WIS values for the
  two methods. This latter choice would penalize a method less for
  doing poorly where the baseline forecaster also does poorly.
  Using instead the geometric mean makes the order of aggregation and
  scaling immaterial since the ratio of geometric means is the same as
  the geometric mean of ratios.
\item If one imagines that a forecaster's WIS is composed of
  multiplicative space-time effects $S_{\ell,t}$ shared across all forecasters,
  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{f,t}$ a
  forecaster-specific error, then
  taking the ratio of two forecasters' geometric mean WIS values will
  effectively cancel these space-time effects.
\end{enumerate}

@@

\section{Statistical significance}

In the Introduction of the main paper, we give some reasons that we avoid making
formal statements about statistical significance, preferring instead to examine
the stability of our results in different contexts. There are strong reasons to
avoid model-based significance tests because the necessary assumptions about
stationarity, independence, and the  model being true (or at least approximately
true) are certainly violated. With those caveats in mind, we undertake two
relatively assumption-lean investigations in this section. The first is a
sign-test for whether the difference between the AR model’s relative WIS  and
each other model’s relative WIS is centered at zero. By ``relative WIS'' we mean
scaled by the strawman as displayed in Figure 3. To mitigate the dependence
across time (which intuitively seems to matter more than that across space), we
computed these tests in a stratified way, where for each forecast date we run a
sign test on the scaled errors between two models over all 440 counties. The
results are plotted as histograms in Figure \ref{fig:sign-test}. In this case,
we use the total relative WIS over all aheads, but the histograms are largely
similar for individual target horizons. If there were no difference, we would
expect to see a uniform distribution. However, for all indicator-assisted
models, we see many more small $P$-values than would be expected if the null
hypothesis (that the AR model is better) were true.

Another relatively assumption-lean method of testing for differences in forecast
accuracy is the Diebold-Mariano test \cite{Diebold:2002, Diebold:2015,
  Harvey:1997}. Essentially, the differences between forecast losses are assumed
to have a constant mean and a covariance that depends on time. Under these
conditions, the asymptotic distribution for the standardized mean of the
differences is limiting normal provided that a heteroskedasticity and
autocorrelation robust estimate of the variance is used. Using the loss as
weighted interval score across all HRRs and horizons (7 to 21 days ahead), we
perform the DM test using both the mean relative to the strawman (as reported in
the manuscript) and the geometric mean relative to the strawman as described
above. The first two rows of Table \ref{tab:dm-test} displays $P$-values for the
test that the indicator-assisted model is no-better than the AR model. Only the
CHNG-CLI model's $P$-value exceeds conventional statistical significance
thresholds.

\section{Bootstrap results}

@@

\section{Leadingness and laggingness}

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness at location $\ell$, we use the cross correlation
function (CCF) 
between the two time series. The $\CCF_\ell(a)$ of an indicator $X_{\ell}$ and case
rates $Y_{\ell}$ is defined as
% \[
% \CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
%   \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
%   \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
% \]
% where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
% shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
% are the sample mean and
% standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
their Pearson correlation where $X_\ell$ has been
% sequence of Pearson correlations for each selected $a$, where we
aligned
with the values of $Y_{\ell}$ that occurred $a$ days earlier.
Thus, for any $a>0$, $\CCF_{\ell}(a)>0$ indicates that
$Y_{\ell,t}$ is moving together with $X_{\ell,t+a}$.
% days in the future.
In this case we say that $X_{\ell}$ is lagging
$Y_{\ell}$. For $a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is
positively correlated with $X_{\ell,t-a}$, so we say that
$X_{\ell}$ leads $Y_{\ell}$.

Figure \ref{fig:ccf-dv-finalized} shows the standardized signals for the HRR
containing Charlotte, North Carolina, from August 1, 2020 until the end of
September. These are the same signals shown in Figure 1 in the manuscript but
using finalized data. To
define ``leadingness'' we compute $\CCF_{\ell}(a)$ (as implemented with the R
function {\tt ccf()}) for each $a\in
\{-15,\ldots,15\}$ using the 56 days leading up to the target date. This is the same
amount of data used to train the forecasters: 21 days of training data, 21 days
to get the response at $a=21$, and 14 days for the longest lagged value. The
orange dashed horizontal line represents the 95\% significance threshold for
correlations based on 56 observations. Any correlations larger in magnitude than
this value are considered statistically significant under the null hypothesis of
no relationship. We define leadingness to be the sum of the significant
correlations that are leading (those above the dashed line with $a<0$) while
laggingness is the same but for $a>0$. In the figure, there are three
significant correlations on the ``leading'' side (at $a = -5, -4, -3$), so
leadingness will be the sum of those values while laggingness is 0:
%correlations, so both scores will be 0:
on September 28 in Charlotte, \dv~is leading cases
leading but not lagging.

@@

\section{Examining data in 2021}

In this section, we investigate the sensitivity of the results to the
period over which we train and evaluate the models.  In the main
paper, we end all evaluation on December 31, 2020.  Figures
\ref{fig:fcast-alldates}--\ref{fig:hot-alldates} show how the
results would differ if we extended this analysis through March
  31, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
the main paper, one sees that as ahead increases most methods now improve
relative to the baseline forecaster. When compared to other methods, 
\chngcli~appears much better than
it had previously; however, all forecasters other than \chngcov~and
\dv~are performing less well relative to the baseline than before.
These changes are likely due to the differing nature of the pandemic
in 2021, with flat and downward trends much more common than upward
trajectories.  Indeed, the nature of the hotspot prediction problem is
quite different in this period.  With a 21-day training window, it is
common for there to be many fewer hotspots in training.

\section{Examining accumulated error over time and space}

>>>>

\section{Finalized Versus Vintage Data}

@@

\section{Aggregation with Geometric Mean}

In this section, we consider using the geometric mean instead of the arithmetic
mean when aggregating the weighted interval score (WIS) across location-time
pairs.  There are three reasons why using the geometric mean may be desirable.
\begin{enumerate}
\item WIS is right-skewed, being bounded below by zero and having occasional 
  very large values.  Figure \ref{fig:wis-densities} illustrates that the
  densities appear roughly log-Gaussian.  The geometric mean is a natural choice  
  in such a context since the relative ordering of forecasters is determined by
  the arithmetic mean of the {\em logarithm} of their WIS values.
\item In the main paper, we report the ratio of the mean WIS of a forecaster to
  the mean WIS of the baseline forecaster. Another choice could be to take the
  mean of the ratio of WIS values for the two methods. This latter choice would
  penalize a method less for doing poorly where the baseline forecaster also
  does poorly.  Using instead the geometric mean makes the order of aggregation
  and scaling immaterial since the ratio of geometric means is the same as the 
  geometric mean of ratios.
\item If one imagines that a forecaster's WIS is composed of multiplicative
  space-time effects $S_{\ell,t}$ shared across all forecasters,
  i.e., $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{f,t}$ a
  forecaster-specific error, then taking the ratio of two forecasters' geometric
  mean WIS values will effectively cancel these space-time effects. 
\end{enumerate}

@@

\section{Statistical Significance}

In the Introduction of the main paper, we give some reasons that we avoid making
formal statements about statistical significance, preferring instead to examine
the stability of our results in different contexts. There are strong reasons to
avoid model-based significance tests because the necessary assumptions about
stationarity, independence, and the  model being true (or at least approximately
true) are certainly violated. With those caveats in mind, we undertake two
relatively assumption-lean investigations in this section. The first is a
sign-test for whether the difference between the AR model’s relative WIS  and
each other model’s relative WIS is centered at zero. By ``relative WIS'' we mean
scaled by the strawman as displayed in Figure 3. To mitigate the dependence
across time (which intuitively seems to matter more than that across space), we
computed these tests in a stratified way, where for each forecast date we run a
sign test on the scaled errors between two models over all 440 counties. The
results are plotted as histograms in Figure \ref{fig:sign-test}. In this case,
we use the total relative WIS over all aheads, but the histograms are largely
similar for individual target horizons. If there were no difference, we would
expect to see a uniform distribution. However, for all indicator-assisted
models, we see many more small p-values than would be expected if the null
hypothesis (that the AR model is better) were true.

Another relatively assumption-lean method of testing for differences in forecast
accuracy is the Diebold-Mariano test \cite{Diebold:2002, Diebold:2015,
Harvey:1997}. Essentially, the differences between forecast losses are assumed
to have a constant mean and a covariance that depends on time. Under these
conditions, the asymptotic distribution for the standardized mean of the
differences is limiting normal provided that a heteroskedasticity and
autocorrelation robust estimate of the variance is used. Using the loss as
weighted interval score across all HRRs and horizons (7 to 21 days ahead), we
perform the DM test using both the mean relative to the strawman (as reported in
the manuscript) and the geometric mean relative to the strawman as described
above. The first two rows of Table \ref{tab:dm-test} displays p-values for the
test that the indicator-assisted model is no-better than the AR model. Only the
CHNG-CLI model's p-value exceeds conventional statistical significance
thresholds.

\section{Bootstrap Results}

@@

\section{Leadingness and Laggingness}

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness at location $\ell$, we use the cross correlation
function (CCF) between the two time series. The $\CCF_\ell(a)$ of an indicator
$X_{\ell}$ and case rates $Y_{\ell}$ is defined as
% \[
% \CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
%   \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
%   \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
% \]
% where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
% shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
% are the sample mean and
% standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
their Pearson correlation where $X_\ell$ has been aligned with the values of
$Y_{\ell}$ that occurred $a$ days earlier.  Thus, for any $a>0$,
$\CCF_{\ell}(a)>0$ indicates that $Y_{\ell,t}$ is moving together with
$X_{\ell,t+a}$.  In this case we say that $X_{\ell}$ is lagging $Y_{\ell}$. For
$a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is positively correlated with
$X_{\ell,t-a}$, so we say that $X_{\ell}$ leads $Y_{\ell}$.

Figure \ref{fig:ccf-dv-finalized} shows the standardized signals for the HRR
containing Charlotte, North Carolina, from August 1, 2020 until the end of
September. These are the same signals shown in Figure 1 in the manuscript but
using finalized data. To define ``leadingness'' we compute $\CCF_{\ell}(a)$ (as
implemented with the R function {\tt ccf()}) for each $a\in \{-15,\ldots,15\}$
using the 56 days leading up to the target date. This is the same amount of data
used to train the forecasters: 21 days of training data, 21 days to get the
response at $a=21$, and 14 days for the longest lagged value. The orange dashed
horizontal line represents the 95\% significance threshold for correlations
based on 56 observations. Any correlations larger in magnitude than this value
are considered statistically significant under the null hypothesis of no
relationship. We define leadingness to be the sum of the significant
correlations that are leading (those above the dashed line with $a<0$) while
laggingness is the same but for $a>0$. In the figure, there are three
significant correlations on the ``leading'' side (at $a = -5, -4, -3$), so
leadingness will be the sum of those values while laggingness is 0: on September
28 in Charlotte, \dv~is leading cases leading but not lagging.

@@

% \section{Examining data in 2021}

% In this section, we investigate the sensitivity of the results to the
% period over which we train and evaluate the models.  In the main
% paper, we end all evaluation on December 31, 2020.  Figures
% \ref{fig:fcast-alldates}--\ref{fig:hot-alldates} show how the
% results would differ if we extended this analysis through March
%   31, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
% the main paper, one sees that as ahead increases most methods now improve
% relative to the baseline forecaster. When compared to other methods, 
% \chngcli~appears much better than
% it had previously; however, all forecasters other than \chngcov~and
% \dv~are performing less well relative to the baseline than before.
% These changes are likely due to the differing nature of the pandemic
% in 2021, with flat and downward trends much more common than upward
% trajectories.  Indeed, the nature of the hotspot prediction problem is
% quite different in this period.  With a 21-day training window, it is
% common for there to be many fewer hotspots in training.

\section{Aggregation Over Time and Space}

### 2021-10-05 forecast/paper/supplement.Rmd
Pass over supplement
- Just formatting and copyedits
- Comment out all eval results from 2021 (but I can't seem to
  compile the Rmd because I'm missing the data files)

<!-- Examining data in 2021 -->

```{r fcast-alldates, fig.cap="Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021."}
plotter(fcasts_honest,
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
```

@@

```{r hot-alldates, fig.cap="Area under the curve for hotspot predictions including data in 2021. Performance degrades relative to the period in 2020. However, there are far fewer hotspots during this period as case rates declined in much of the country."}
plotter_hotspots(hotspots_honest)
```

>>>>

<!-- Examining data in 2021 -->
<!--
```{r fcast-alldates, fig.cap="Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021."}
plotter(fcasts_honest,
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
```

@@

```{r hot-alldates, fig.cap="Area under the curve for hotspot predictions including data in 2021. Performance degrades relative to the period in 2020. However, there are far fewer hotspots during this period as case rates declined in much of the country."}
plotter_hotspots(hotspots_honest)
```
-->
